---
title: "Group Mermaid 206 Assignment 4"
author: "Callie Steffen, Vanessa Guenther, Renee Albrecht"
date: "11/12/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Part 1.Describe trends in lobster abundance (counts) and fishing pressure (trap buoys) at the five locations from 2012 - 2017. 

```{r, include=FALSE}
###Loading Necessary Packages
library(tidyverse)
library(effsize)
library(kableExtra)
library(pwr)
library(knitr)
library(plotly)
library(ggplot2)
library(car)
library(onewaytests)
library(vcdExtra)
library(xtable)
install.packages("ggsignif")
library(ggsignif)

###Reading in the data sets and renaming for convenience

size_abund<- read.csv("lobster_size_abundance.csv")
traps<-read.csv("lobster_traps.csv")

##Filtering out the transects from data set and only keeping the year not month(keeping year, site, count)
##Basic data wrangling

lobster_abund<- select(size_abund,YEAR,SITE,COUNT)

traps_site <-select(traps,YEAR,SITE,TRAPS) %>%    ##(keeping year, site, traps)
  filter(SITE=="AQUE"|SITE=="NAPL"|SITE=="MOHK"|SITE=="IVEE"|SITE=="CARP") ###Filtering for the 5 sites we are looking at

###Exploratory Graphs of the Data Frames

#Histogram for Lobster Abundance

abund_histogram<-ggplot(lobster_abund, aes(x=COUNT)) +
  geom_histogram() 

abund_histogram


###QQ plot for Lobster Abundanance
lobster_qq<-ggplot(lobster_abund, aes(sample =COUNT))+
  geom_qq()
lobster_qq

##Histogram for Traps

traps_histogram<-ggplot(traps_site, aes(x=TRAPS))+
  geom_histogram()

traps_histogram

###QQ Plot for Traps
traps_qq<-ggplot(traps_site, aes(sample=TRAPS)) +
  geom_qq()
traps_qq

scatter_abundance<-ggplot(lobster_abund, aes(x=YEAR, y=COUNT))+
  geom_point()+
  facet_wrap(~SITE, scale = "free") +
  theme_classic()

scatter_abundance

scatter_traps<-ggplot(traps_site, aes(x=YEAR, y=TRAPS))+
  geom_point(aes(color="COUNT"))+
  facet_wrap(~SITE, scale="free") +
  scale_y_continuous(name="number of lobsters",limits=c(0,25))+
  theme_classic()

scatter_traps
ggplotly(scatter_traps)

  


```
####Part 2. Compare mean lobster sizes (carapace length (mm)) across the five sites for lobster observations collected in 2017.

```{r}
# Create data frame filter for 5 locations and year 2017, select size (remove all -9999 values and turn data into case-format)

compare_ls <- select(size_abund,YEAR,SITE,SIZE,COUNT) %>% 
  filter(YEAR == 2017, 
         SITE=="AQUE"|SITE=="NAPL"|SITE=="MOHK"|SITE=="IVEE"|SITE=="CARP",
         SIZE!= "-99999")
#Remove multiple counts and organize into rows 
ls_comp <- uncount(compare_ls, weights = COUNT, .remove = TRUE, .id = NULL)
#Count in decimal numbers ? 


counts <- with(ls_comp, table(YEAR,SITE,SIZE))
View(counts)
View (ls_comp)

```


```{r}
# Exploratory graphs 

hist_ls <- ggplot(ls_comp, aes(x = SIZE))+
  geom_histogram(bins = 10) +
  facet_wrap(~ SITE, scale = "free")

hist_ls

qq_ls <-ggplot(ls_comp, aes(sample = SIZE)) +
  geom_qq(bins = 10) +
  facet_wrap(~ SITE, scale = "free")

qq_ls

# Groups approx. normally distributed  

box_ls <-  ggplot(ls_comp, aes(x = SITE, y = SIZE)) +
  geom_boxplot(width = .4) +
  geom_jitter(width = 0.1, alpha = 0.5, aes(color = SITE))

box_ls


```

```{r}
#Leven's Test 
variances_ls <- ls_comp %>% 
  group_by(SITE) %>% 
  summarize(
    mean = mean(SIZE),
    sd = sd(SIZE),
    variance = var(SIZE)
  )
View(variances_ls)

# assume equal variance 

levene_ls <- leveneTest(SIZE ~ SITE, data = ls_comp)
levene_ls

#Ho: No difference in variances (Variances are equal)
# Ha: Variances are NOT equal

# There issignificant differences in variances between SITES ? (or use robust SE) 16  Can we still use anova if L-test not affected ? or Welsh ?  

#ANOVA One-Way 
#H0: All means are equall 
#HA: At least two mean are not equal 

ls_anova <- aov(SIZE ~ SITE, data = ls_comp)
ls_sum <- summary(ls_anova)
ls_sum
#At least two means are NOT equal ?

# Which ones ? 
ls_tukey <- TukeyHSD(ls_anova)
ls_tukey


# Table comparing ??

stats_ls <- ls_comp %>% 
  group_by(SITE) %>% 
  summarize(
    mean = round(mean(SIZE),2),
    sd = round(sd(SIZE),2))


figure_ls <- ggplot(stats_ls, aes(x = SITE, y = mean)) +
  geom_col(colour = NA, fill = "gray50", width = 0.5) +
  geom_errorbar(aes(ymax = mean + sd, ymin = mean - sd), width = 0.1)+
   scale_y_continuous(expand = c(0,0), limits = c(0,95))+
    labs(y=expression(Mean~Lobster~Size~~(m)))+
  scale_x_discrete() +
  xlab("\nResearch Sites")+
   geom_signif() #  GEOM SIGNIF ?????


figure_ls

```

### Part 3: Comparing changes in lobster size at MPA and non-MPA sites (comparing only 2012 and 2017 sizes)
At Isla Vista and Naples Reef, the two protected MPA sites (with zero fishing pressure), how do lobster sizes
in 2012 and 2017 compare? At the non-MPA sites?


```{r}
#create dataframe to compare lobster size in 2012 and 2017 at MPA and non-MPA sites 

MPA_df <- size_abund %>%
  filter(YEAR == "2012" | YEAR == "2017",  SIZE != "-99999") %>%
  mutate(
    MPA_STATUS = case_when( 
      SITE == "IVEE" ~ "MPA",
      SITE == "NAPL" ~ "MPA",
      SITE == "AQUE" ~ "NON-MPA",
      SITE == "CARP" ~ "NON-MPA",
      SITE == "MOHK" ~ "NON-MPA")) %>%
  select(YEAR, SITE, SIZE, COUNT, MPA_STATUS)

#arrange data so that each lobster has its own row 
MPAdf <- as.data.frame(expand.dft(MPA_df, freq = "COUNT"))
     

View(MPAdf)

#create data summary frame of lobster size median, max, mean, SD, and sample size 

MPA_table <- as.data.frame(expand.dft(size_abund, freq = "COUNT")) %>%
group_by(YEAR) %>% #group data together by same month
  summarize(
    median = round(median(SIZE),2), #find median, round to 2 digits
    max = max(SIZE), #identify maximum 
    mean = round(mean(SIZE),2), #find mean, round to 2 digits 
    stdev = round(sd(SIZE),2), #find SD, round to 2 digits
    length(SIZE)) #find sample size 

View(MPA_table)

```


```{r}


#Create exloratory histograms and qq plots

#histogram to compare 2012 and 2017 lobster sizes 
MPAhist <- ggplot(MPAdf, aes(x = SIZE)) +
  geom_histogram(aes(bins=15, fill = SITE)) +
                   facet_wrap(~YEAR ~MPA_STATUS, scale = "free") +
                   theme_classic()
MPAhist

#qq plot to compare 2012 and 2017 lobster sizes 
MPAqq <- ggplot(MPAdf, aes(sample = SIZE)) +
  geom_qq(aes(color = SITE)) + 
  facet_wrap(~YEAR ~MPA_STATUS, scale = "free") + 
  theme_classic()

MPAqq




```


## Is there a significant difference between lobster size at MPA sites in 2012 and 2017? 

## Is there a significant difference between lobster size at non-MPA sites in 2012 and 2017? 

## Is there a significant difference between lobster size at MPA vs. Non-MPA sites in 2017? 




