---
title: "Group Mermaid 206 Assignment 4"
author: "Callie Steffen, Vanessa Guenther, Renee Albrecht"
date: "11/12/2018"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Part 1.Describe trends in lobster abundance (counts) and fishing pressure (trap buoys) at the five locations from 2012 - 2017. 

```{r, include=FALSE}
###Loading Necessary Packages
library(tidyverse)
library(effsize)
library(kableExtra)
library(pwr)
library(knitr)
library(plotly)
library(ggplot2)
library(car)
library(onewaytests)
library(vcdExtra)
library(xtable)
library(ggsignif)

###Reading in the data sets and renaming for convenience

size_abund<- read.csv("lobster_size_abundance.csv")
traps<-read.csv("lobster_traps.csv")

##Filtering out the transects from data set and only keeping the year not month(keeping year, site, count)
##Basic data wrangling

lobster_abund<- select(size_abund,YEAR,SITE,COUNT)

traps_site <-select(traps,YEAR,SITE,TRAPS) %>%    ##(keeping year, site, traps)
  filter(SITE=="AQUE"|SITE=="NAPL"|SITE=="MOHK"|SITE=="IVEE"|SITE=="CARP") ###Filtering for the 5 sites we are looking at

###Exploratory Graphs of the Data Frames

#Histogram for Lobster Abundance

abund_histogram<-ggplot(lobster_abund, aes(x=COUNT)) +
  geom_histogram() 

abund_histogram


###QQ plot for Lobster Abundanance
lobster_qq<-ggplot(lobster_abund, aes(sample =COUNT))+
  geom_qq()
lobster_qq

##Histogram for Traps

traps_histogram<-ggplot(traps_site, aes(x=TRAPS))+
  geom_histogram()

traps_histogram

###QQ Plot for Traps
traps_qq<-ggplot(traps_site, aes(sample=TRAPS)) +
  geom_qq()
traps_qq

scatter_abundance<-ggplot(lobster_abund, aes(x=YEAR, y=COUNT))+
  geom_point()+
  facet_wrap(~SITE, scale = "free") +
  theme_classic()

scatter_abundance

scatter_traps<-ggplot(traps_site, aes(x=YEAR, y=TRAPS))+
  geom_point(aes(color="COUNT"))+
  facet_wrap(~SITE, scale="free") +
  scale_y_continuous(name="number of lobsters",limits=c(0,25))+
  theme_classic()

scatter_traps
ggplotly(scatter_traps)

  


```
####Part 2. Compare mean lobster sizes (carapace length (mm)) across the five sites for lobster observations collected in 2017.

```{r}
# Create data frame filter for 5 locations and year 2017, select size (remove all -9999 values and turn data into case-format)

compare_ls <- select(size_abund,YEAR,SITE,SIZE,COUNT) %>% 
  filter(YEAR == 2017, 
         SITE=="AQUE"|SITE=="NAPL"|SITE=="MOHK"|SITE=="IVEE"|SITE=="CARP",
         SIZE!= "-99999")
#Remove multiple counts and organize into rows 
ls_comp <- uncount(compare_ls, weights = COUNT, .remove = TRUE, .id = NULL)
#Count in decimal numbers ? 

#View(counts)
View (ls_comp)

```


```{r}
# Exploratory graphs 

hist_ls <- ggplot(ls_comp, aes(x = SIZE))+
  geom_histogram(bins = 10) +
  facet_wrap(~ SITE, scale = "free")

hist_ls

qq_ls <-ggplot(ls_comp, aes(sample = SIZE)) +
  geom_qq(bins = 10) +
  facet_wrap(~ SITE, scale = "free")

qq_ls

# Groups approx. normally distributed  

box_ls <-  ggplot(ls_comp, aes(x = SITE, y = SIZE)) +
  geom_boxplot(width = .4) +
  geom_jitter(width = 0.1, alpha = 0.5, aes(color = SITE))

box_ls


```

```{r}
#Leven's Test 
variances_ls <- ls_comp %>% 
  group_by(SITE) %>% 
  summarize(
    mean = mean(SIZE),
    sd = sd(SIZE),
    variance = var(SIZE)
  )
View(variances_ls)

# assume equal variance 

levene_ls <- leveneTest(SIZE ~ SITE, data = ls_comp)
levene_ls

#Ho: No difference in variances (Variances are equal)
# Ha: Variances are NOT equal

# There issignificant differences in variances between SITES ? (or use robust SE) 16  Can we still use anova if L-test not affected ? or Welsh ?  

#ANOVA One-Way 
#H0: All means are equall 
#HA: At least two mean are not equal 

ls_anova <- aov(SIZE ~ SITE, data = ls_comp)
ls_sum <- summary(ls_anova)
ls_sum
#At least two means are NOT equal ?

# Which ones ? 
ls_tukey <- TukeyHSD(ls_anova)
ls_tukey


# Table comparing ??

stats_ls <- ls_comp %>% 
  group_by(SITE) %>% 
  summarize(
    mean = round(mean(SIZE),2),
    sd = round(sd(SIZE),2))


figure_ls <- ggplot(stats_ls, aes(x = SITE, y = mean)) +
  geom_col(colour = NA, fill = "#FF6666", width = 0.5) +
  geom_errorbar(aes(ymax = mean + sd, ymin = mean - sd), width = 0.1)+
   scale_y_continuous(expand = c(0,0), limits = c(0,120))+
    labs(y=expression(Mean~Lobster~Size~~(m)))+
  scale_x_discrete() +
  xlab("\nResearch Sites")+
   geom_signif(comparisons = list(c("NAPL","CARP")), annotations = "p = 0.023", y_position = 95, tip_length = 0.1, size = 0.5, textsize = 3)+
  geom_signif(comparisons = list(c("NAPL","IVEE")), annotations = "p = 0.004", y_position = 110, tip_length = 0.1, size = 0.5, textsize = 3)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),plot.title = element_text(face ="bold"))+
  ggtitle("Lobster Size Distribution at SBC LTER Sites")
  
 
 
figure_ls

#numer of samples for each site
ls_comp %>% count(SITE)

```

**Figure XX. Lobsrer Size Destribution at SBC LTER SITES.** Mean lobster size(mm) for locations AQUE (n=67), CARP (n=705), IVEE (n=606), MOHK (n=178), and NAPL (n=122) Santa Barbara County. Error bars indicate +/- 1 standard deviation. Brackets indicate *p*-values for significantly different means as determined by one-way ANOVA with Tukey's HSD (F(3,42) = ???, *p* < 0.001, with $\alpha$ = 0.05 throughout). Data source: Santa Barbara Coastal (bc.lternet.edu/cgi-bin/showDataset.cgi?docid=knb-lter-sbc.77)






### Part 3: Comparing changes in lobster size at MPA and non-MPA sites (comparing only 2012 and 2017 sizes)
At Isla Vista and Naples Reef, the two protected MPA sites (with zero fishing pressure), how do lobster sizes
in 2012 and 2017 compare? At the non-MPA sites?


```{r}
#create dataframe to compare lobster size in 2012 and 2017 at MPA and non-MPA sites 

MPA_df <- size_abund %>%
  filter(YEAR == "2012" | YEAR == "2017",  SIZE != "-99999") %>%
  mutate(
    MPA_STATUS = case_when( 
      SITE == "IVEE" ~ "MPA",
      SITE == "NAPL" ~ "MPA",
      SITE == "AQUE" ~ "NON-MPA",
      SITE == "CARP" ~ "NON-MPA",
      SITE == "MOHK" ~ "NON-MPA")) %>%
  select(YEAR, SITE, SIZE, COUNT, MPA_STATUS)

#arrange data so that each lobster has its own row 
MPAdf <- as.data.frame(expand.dft(MPA_df, freq = "COUNT"))
     
#View(MPAdf)

#create data summary frame of lobster size median, max, mean, SD, and sample size 

MPA_table <- as.data.frame(expand.dft(size_abund, freq = "COUNT")) %>%
group_by(YEAR) %>% #group data together by same month
  summarize(
    median = round(median(SIZE),2), #find median, round to 2 digits
    max = max(SIZE), #identify maximum 
    mean = round(mean(SIZE),2), #find mean, round to 2 digits 
    stdev = round(sd(SIZE),2), #find SD, round to 2 digits
    length(SIZE)) #find sample size 

#View(MPA_table)

```


```{r}


#Create exloratory histograms and qq plots

#histogram to compare 2012 and 2017 lobster sizes 
MPAhist <- ggplot(MPAdf, aes(x = SIZE)) +
  geom_histogram(aes(bins=15, fill = SITE)) +
                   facet_wrap(~YEAR ~MPA_STATUS, scale = "free") +
                   theme_classic()
MPAhist

#qq plot to compare 2012 and 2017 lobster sizes 
MPAqq <- ggplot(MPAdf, aes(sample = SIZE)) +
  geom_qq(aes(color = SITE)) + 
  facet_wrap(~YEAR ~MPA_STATUS, scale = "free") + 
  theme_classic()

MPAqq




```

## Part 3 question of interest: 
## Is there a significant difference between lobster size at each of the sites in 2012 and 2017? 


```{r}

#create data frames to perform statistical tests to answer the question above 

#AQUE dataframe
AQUEdf <- MPAdf %>%
  filter(SITE == "AQUE")
#View(AQUEdf)

#CARP datafram
CARPdf <- MPAdf %>%
  filter(SITE == "CARP")
#View(CARPdf)

#MOHK dataframe
MOHKdf <- MPAdf %>%
  filter(SITE == "MOHK")
#View(MOHKdf)

#IVEE dataframe
IVEEdf <- MPAdf %>%
  filter(SITE == "IVEE")
#View(IVEEdf)

#NAPL dataframe 
NAPLdf <- MPAdf %>%
  filter(SITE == "NAPL")
#View(NAPLdf)

########################

mpa_sites <- MPAdf %>%
  filter(MPA_STATUS == "MPA")

#View(mpa_sites)

nonmpa_sites <- MPAdf %>%
  filter(MPA_STATUS == "NON-MPA")

#View(nonmpa_sites)

post_sites <- MPAdf %>%
  filter(YEAR == "2017")

#View(post_sites)

```

```{r}

<<<<<<< HEAD

legal_lobster<-ls_comp %>% 
  filter(YEAR=="2017",SIZE !="NA") %>% 
  mutate(legality= ifelse(SIZE < 82.6, "ILLEGAL", "LEGAL")) %>% 
  select(SITE,legality) %>% 
  count(SITE,legality) %>% 
  spread(legality,n) %>% 
  select(-SITE)

rownames(legal_lobster)<- c("AQUE", "CARP", "IVEE","MOHK","NAPL")

prop_lobster<-prop.table(as.matrix(legal_lobster),1)

prop_lobster

chi_lobster<- chisq.test(legal_lobster)
chi_lobster
=======
#AQUE site (comparing 2012 and 2017 sizes)

#F-Test for equal variances 
#AQUE_ftest <- AQUEdf %>%
  #var.test(SIZE ~ YEAR, data = .)

#AQUE_ftest 
#RESULT: Variances are NOT equal 

#T-test 
AQUE_ttest <- AQUEdf %>%
  t.test(SIZE ~ YEAR, data =.)

#AQUE_ttest

#NOT significantly different (p-value = 0.1907)

```

```{r}

#CARP site (comparing 2012 and 2017 sizes)

#F-Test for equal variances 
#CARP_ftest <- CARPdf %>%
  #var.test(SIZE ~ YEAR, data = .)

#CARP_ftest 
#RESULT: Variances are NOT equal  

#T-test 
CARP_ttest <- CARPdf %>%
  t.test(SIZE ~ YEAR, data =.)

#CARP_ttest

#NOT significantly different (p-value = 0.2211)

```

```{r}

#MOHK site (comparing 2012 and 2017 sizes)

#F-Test for equal variances 
#MOHK_ftest <- MOHKdf %>%
  #var.test(SIZE ~ YEAR, data = .)

#MOHK_ftest 
#RESULT: Variances are NOT equal  

#T-test 
MOHK_ttest <- MOHKdf %>%
  t.test(SIZE ~ YEAR, data =.)

#MOHK_ttest

#SIGNIFICANTLY DIFFERENT (p-value = 0.0001599)

```

```{r}
#IVEE site (comparing 2012 and 2017 sizes)

#F-Test for equal variances 
#IVEE_ftest <- IVEEdf %>%
  #var.test(SIZE ~ YEAR, data = .)

#IVEE_ftest 
#RESULT: Variances are NOT equal  

#T-test 
IVEE_ttest <- IVEEdf %>%
  t.test(SIZE ~ YEAR, data =.)

#IVEE_ttest

#SIGNIFICANTLY DIFFERENT (p-value = 0.0361)



```

```{r}

#NAPL site (comparing 2012 and 2017 sizes)

#F-Test for equal variances 
#NAPL_ftest <- NAPLdf %>%
  #var.test(SIZE ~ YEAR, data = .)

#NAPL_ftest 
#RESULT: Variances are NOT equal  

#T-test 
NAPL_ttest <- NAPLdf %>%
  t.test(SIZE ~ YEAR, data =.)

#NAPL_ttest

#NOT significantly different (p-value = 0.5373)


```

```{r}

#Cohen's D

IVEE_effect <- cohen.d(SIZE ~ YEAR, IVEEdf)
IVEE_effect

## Result: -0.3775177 (small)

MOHK_effect <- cohen.d(SIZE ~ YEAR, MOHKdf)
MOHK_effect

## Result: 0.5408116 (medium)

```


```{r}
####
#Comparing lobster size at ALL MPA sites (IVEE and NAPL) in 2012 and 2017
####

#F Test for equal variance 
#H0: Ratio of variances is = 1 
#HA: Ratio of variances is NOT = 1

mpa_ftest <- mpa_sites %>%
  var.test(SIZE ~ YEAR, data = .)

mpa_ftest
# Test determined variances are NOT equal 

#Two sample two sided T-Test
#H0: The difference between the means = 0
#HA: The difference between the means is NOT = 0

mpa_ttest <- mpa_sites %>%
  t.test(SIZE ~ YEAR, data =., var.equal = TRUE)

mpa_ttest

#NOT SIGNIFICANTLY DIFFERENT (p-value = 0.05576)

```

```{r}
###
#Comparing lobster size at ALL non-MPA sites (CARP, MOHK, and AQUE) in 2012 and 2017
###

#F Test for equal variance 
#H0: Ratio of variances is = 1 
#HA: Ratio of variances is NOT = 1

nonmpa_ftest <- nonmpa_sites %>%
  var.test(SIZE ~ YEAR, data = .)

nonmpa_ftest
# Test determined variances are NOT equal 

#Two sample two sided T-Test
#H0: The difference between the means = 0
#HA: The difference between the means is NOT = 0

nonmpa_ttest <- nonmpa_sites %>%
  t.test(SIZE ~ YEAR, data =., var.equal = TRUE)

nonmpa_ttest

#SIGNIFICANTLY DIFFERENT (p-value = 0.007093)

```

```{r}
###
#Comparing lobster size at MPA vs. Non-MPA sites in 2017
###


```

```{r}

#Create finalized data table 

sum_df <- MPAdf %>%
  mutate(
    SITE_FULL = case_when( 
      SITE == "IVEE" ~ "Isla Vista (IVEE)",
      SITE == "NAPL" ~ "Naples Reef (NAPL)",
      SITE == "AQUE" ~ "Arroyo Quemado (AQUE)",
      SITE == "CARP" ~ "Carpinteria (CARP)",
      SITE == "MOHK" ~ "Mohawk Reef (MOHK)"))

sizesum <- sum_df %>%
  group_by(MPA_STATUS, SITE_FULL, YEAR) %>%
  summarize(
    mean = round(mean(SIZE), 1),
    stdev = round(sd(SIZE), 1),
    length(SIZE))
  
sizesum
>>>>>>> 03bcfded3e96b85c74eb390e1ff33daa0c9b9341



```

```{r}

kable(sizesum, col.names = c("MPA Status", "Site", "Year", "Mean (mm)", "SD (mm)", "sample size"), caption = "**Table 1**. California spiny lobster (*Panulirus interruptus*) carapace length (mm) at five Santa Barbara Channel reefs in 2012 and 2017. Source: Santa Barbara Coastal Long Term Ecological Research") %>%
  kable_styling(bootstrap_options = c("striped"), full_width = F) %>%
  collapse_rows()



```

